---
title: "2020 Census response rates API"
author: Bernardo Lopez Vicencio, [NewsCounts](https://newscounts.org/)
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```


The Census Bureau, starting on March 20, has been releasing daily data of the [2020 Census response rates](https://2020census.gov/en/response-rates.html) through a well documented [API](https://www.census.gov/data/developers/data-sets/decennial-response-rates.html). In NewsCounts we have been keeping track of these daily data and now we have it available for everyone who wants to use it through our own API.

Our API offers daily snapshots of the response rates, as opposed to the cumulative data that the Bureau’s API provides. Also our API gives the ability of retrieving data in JSON format, CSV files that can be opened in Excel, and it can create automated plots of the response rates.

## Use examples

Downloading the data is as simple as going to a specific URL in your favorite web browser. For example, to get the response rates for all states go to:

```
https://newscounts.appspot.com/api/response_rates/state?FORMAT=CSV
```

```{r, include=FALSE}
state_df <- read_csv("https://newscounts.appspot.com/api/response_rates/state?FORMAT=CSV")

county_df <- read_csv("https://newscounts.appspot.com/api/response_rates/county?FORMAT=CSV")
```

With the daily state response data we can find out, for example, which New York counties have higher response rates and which counties are doing worst by the date this document is created (March 29).

* Counties with higher response rate

```{r, echo=FALSE}
new_york <-county_df %>%
  filter(county_df$state_short =="NY", county_df$RESP_DATE == "2020-03-29") %>%
  mutate(CRRALL=as.numeric(as.character(CRRALL)))

best_ny <- new_york %>%
  arrange(desc(CRRALL))

knitr::kable(best_ny[0:5,c("county_name","CRRALL")])
```

* Counties with low response rate

```{r, echo=FALSE}
worst_ny <- new_york %>%
  arrange(CRRALL)

knitr::kable(worst_ny[0:5,c("county_name","CRRALL")])
```


By having the daily data we can do more interesting things, such as time series to look for trends in the response rates. We have included a function in the API that gives you automatic time series plots.
For this example, by using the following URL, I can get a plot that compares the cumulative response rate for the states of California, New York, Pennsylvania and Alaska.

```{r}
image_url <- "https://newscounts.appspot.com/api/response_rates/state?FORMAT=PNG&DATA=CRRALL&STATE=CA,NY,PA,AK"
```

![Example automatic plot.](`r image_url`)

As days pass by, we can look to more interesting trends not just for the state level, but also for counties and tracts.

## Documentation

### State data
**Method:** GET

**URL:** ```https://newscounts.appspot.com/api/response_rates/state```

**Response format:** CSV and JSON for data, PNG for plot image.

**Parameters:**

|  Name | Required | Description | Example |
|--------|---------|---------|-----------|
|```FORMAT```| Required | JSON for data in that format. CSV if you want to download data in a CSV file. PNG if you want a time series plot. | JSON |
|```DATA``` | Required if ```FORMAT``` is PNG | List of values you want to get. It has to be at least one of the following separated by a comma: CRRAL, CRRINT, DAVG, DINTAVG, DRRALL, DRRINT. Select only one if ```FORMAT``` is PNG. If not present, all of them will be returned. | CRRAL,DRRALL|
|```FROM``` | Required when ```TO``` is present | Date with format YYYY-MM-DD | 2020-03-23|
|```TO``` | Required when ```FROM``` is present | Date with format YYYY-MM-DD | 2020-03-25|
|```STATE```| Required if ```FORMAT``` is PNG | ISO 2 letter abbreviation for U.S. states. You can select many states separated by a comma.  | NY,CA |

**Example request:**

```
https://newscounts.appspot.com/api/response_rates/state?FORMAT=JSON&DATA=CRRALL,DRRALL&STATE=NY&FROM=2020-03-23&TO=2020-03-25
```

**Example response:**

```
{"CRRALL":22.3,"DRRALL":2.2,"GEO_ID":"0400000US36","RESP_DATE":"2020-03-25","state":36,"state_name":"New York","state_short":"NY"},
{"CRRALL":20.1,"DRRALL":2.2,"GEO_ID":"0400000US36","RESP_DATE":"2020-03-24","state":36,"state_name":"New York","state_short":"NY"},
{"CRRALL":17.9,"DRRALL":1.5,"GEO_ID":"0400000US36","RESP_DATE":"2020-03-23","state":36,"state_name":"New York","state_short":"NY"}
```

### County data
**Method:** GET

**URL:** ```https://newscounts.appspot.com/api/response_rates/county```

**Response format:** CSV and JSON for data, PNG for plot image.

**Parameters:**

|  Name | Required | Description | Example |
|--------|---------|---------|-----------|
|```FORMAT```| Required | JSON for data in that format. CSV if you want to download data in a CSV file. PNG if you want a time series plot. | JSON |
|```DATA``` | Required if ```FORMAT``` is PNG | List of values you want to get. It has to be at least one of the following separated by a comma: CRRAL, CRRINT, DAVG, DINTAVG, DRRALL, DRRINT. Select only one if ```FORMAT``` is PNG. If not present, all of them will be returned. | CRRAL,DRRALL|
|```FROM``` | Required when ```TO``` is present | Date with format YYYY-MM-DD | 2020-03-23|
|```TO``` | Required when ```FROM``` is present | Date with format YYYY-MM-DD | 2020-03-25|
|```STATE```| Required if ```COUNTY``` is present | ISO 2 letter abbreviation for U.S. states | NY |
|```COUNTY```| Required if ```FORMAT``` is PNG | [FIPS county codes](https://www.census.gov/prod/techdoc/cbp/cbp95/st-cnty.pdf). You can select many counties separated by a comma. | 1,3 |

**Example request:**

```
https://newscounts.appspot.com/api/response_rates/county?FORMAT=JSON&DATA=CRRALL,DRRALL&STATE=NY&COUNTY=3&FROM=2020-03-23&TO=2020-03-25
```

**Example response:**

```
{"CRRALL":19.6,"DRRALL":1.5,"GEO_ID":"0500000US36003","RESP_DATE":"2020-03-25","county":"003","county_name":"Allegany","state":36,"state_name":"New York","state_short":"NY"},
{"CRRALL":18.1,"DRRALL":1.0,"GEO_ID":"0500000US36003","RESP_DATE":"2020-03-24","county":"003","county_name":"Allegany","state":36,"state_name":"New York","state_short":"NY"},
{"CRRALL":17.1,"DRRALL":0.5,"GEO_ID":"0500000US36003","RESP_DATE":"2020-03-23","county":"003","county_name":"Allegany","state":36,"state_name":"New York","state_short":"NY"}
```

### Tract data
**Method:** GET

**URL:** ```https://newscounts.appspot.com/api/response_rates/tract```

**Response format:** CSV and JSON for data, PNG for plot image.

**Parameters:**

|  Name | Required | Description | Example |
|--------|---------|---------|-----------|
|```FORMAT```| Required | JSON for data in that format. CSV if you want to download data in a CSV file. PNG if you want a time series plot. | JSON |
|```DATA``` | Required if ```FORMAT``` is PNG | List of values you want to get. It has to be at least one of the following separated by a comma: CRRAL, CRRINT, DAVG, DINTAVG, DRRALL, DRRINT. Select only one if ```FORMAT``` is PNG. If not present, all of them will be returned. | CRRAL,DRRALL|
|```FROM``` | Required when ```TO``` is present | Date with format YYYY-MM-DD | 2020-03-23|
|```TO``` | Required when ```FROM``` is present | Date with format YYYY-MM-DD | 2020-03-25|
|```STATE```| Required when ```COUNTY``` is present | ISO 2 letter abbreviation for U.S. states | NY |
|```COUNTY```| Required when ```TRACT``` is present | [FIPS county codes](https://www.census.gov/prod/techdoc/cbp/cbp95/st-cnty.pdf) | 1 (For Autauga County in Alabama) |
|```TRACT``` | Required if ```FORMAT``` is PNG | 2020 Census tract number. You can select many tracts separated by a comma. | 020100|


**Example request:**

```
https://newscounts.appspot.com/api/response_rates/tract?FORMAT=JSON&DATA=CRRALL,DRRALL&STATE=NY&COUNTY=3&FROM=2020-03-25&TO=2020-03-25&TRACT=950600
```

**Example response:**

```
{"CRRALL":24.9,"DRRALL":2.2,"GEO_ID":"1400000US36003950600","RESP_DATE":"2020-03-25","county":"003","county_name":"Allegany","state":36,"state_name":"New York","state_short":"NY","tract":"950600"}
```

## DATA labels definition
* **CRRALL:** Cumulative Self-Response Rate - Overall
* **CRRINT:** Cumulative Self-Response Rate - Internet
* **DAVG:** Average Daily Overall Self-Response Rate
* **DINTAVG:** Average Daily Internet Self-Response Rate
* **DRRALL:** Daily Self-Response Rate - Overall
* **DRRINT:** Daily Self-Response Rate - Internet


